%%
%% Template conclusion.tex
%%

\chapter{Conclusion}
\label{cha:conc}

In this chapter we will outline a summary of the work completed during this thesis and discuss our proposal for future work in this area.

\section{Summary}
\label{sec:conc}

In this thesis we have tested and compared an exhaustive list of different affinity features across varied exposure sizes.

We have found that \emph{user interaction} affinity features in themselves are not predictive of user likes, however coupled against 
exposure, they show a comprehensive predictive improvement over our baselines.

Conversely, we have shown the interesting result that \emph{user preference} affinity features are more predictive of user likes
and this trend continues across user exposure.

We have investigated individual groups and found the most predictive were highly localised with a medium
user frequency, while the most predictive pages were much more general and of a higher relative user frequency.

To answer the question initially proposed for this thesis, we have shown the affinity features which provide the highest predictiveness of user likes come from \emph{user preferences} and not
\emph{user interactions}. The most predictive features found in this analysis are \emph{favourites}, \emph{group memberships} 
and \emph{page likes}, additionally a combination of these individually predictive features results in the most predictive classification.

This provides the exciting and novel insights examined during this thesis.

\section{Future Work}
\label{sec:ftw}

Proposed future work can be summarised under the following points:

\begin{itemize}
\item \textbf{Increase size ranges}: Given our maximum test sizes used for \emph{groups} and \emph{pages} and the trend that an increased
size range resulted in better prediction, this size could be 
increased to find the optimal testing range for each of our classifiers.
\item \textbf{Passive likes}: Given the Facebook model of allowing users to like but not dislike data, explicit dislike data can not be 
gleaned from Facebook, which is why the active likes sourced from the NICTA app were used during this evaluation. An approach could be developed 
which can predict whether a user will have seen an item (online timestamps, recent interactions with a user, etc) and we can infer that if the 
user did not like the item then they disliked it. This data set could then be applied to the testing methodology outlined in the above chapters.
\item \textbf{Cold start}: Leaving out some subset of users when training our models, but including them during testing to explore their 
effects on results.
\item \textbf{General user set}: Such as the study done by \cite{jugand} which comprised of the entire active social network of 721 million users 
(as of May 2011), applying these methods to a data set which is more indicative of the general Facebook user population could offer more 
generalisable results.
\item \textbf{Bayesian Model Averaging}: Combining the most predictive qualities of the individual classifiers by sampling and
weighting the results by applying Bayes' law.
\end{itemize}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "thesis"
%%% End: 
