%!TEX root = document.tex

{\bf TODO: need to address all missing info in reviews and probably rewrite for clarity: what the app did, what suggestions it made to user, the fact that it was offline batch and not online, whether there was rank bias -- no, independent ratings.  Tables provide good summaries so should not need to be too verbose in this section.}

We built a Facebook App\footnote{Name and link omitted
for anonymity.} to collect information about users, 
their interactions and preferences on daily recommended links.  
%data about Facebook users and their Facebook friends
%are collected through our Facebook App. 
Our dataset contains information about each App user along with their
interactions with all of their friends.  The
data collection is performed with full permission from the user and in
accordance with an approved Ethics Protocol\footnote{Link omitted for
anonymity.}.

%% I'm going to let the table do the summarizing this time
\eat{
Over 119 users installed the Facebook App during the evaluation
period.\footnote{In order to collect full interactions among App users
as required in our experimentation, our App requested to collect
information posted to \emph{friends'} timelines.  With such expressive
permissions concerning friend interactions, most potential users were
hesitant to install the App hence after a one month user drive period
we were able to attract the pool of users used in the experiments.}
From these core App users, the App has access to their detailed
Facebook profiles and their interactions with a total of 38,259
friends.
}

%interaction data.  
While we have complete interaction data for the App users with their
friends, and profile data (including wall post data) for the App users
and friends, we do not have complete interactions for the App users'
friends (unless they themselves are App users).  Hence in the
forthcoming analysis, we limit our evaluation to App users for which
we are assured to have full interaction data.

Our App tracks app user's (and their friends') details and
interactions on Facebook.  Interactions that occur through wall posts
provide a rich variety of content and interaction data.  We
distinguish four Facebook items from wall posts: general posts (e.g.,
status updates, activity updates such as new friends), links, photos
and videos. Four main interactions on these items are permitted by
Facebook: posting an item to a friend's wall, commenting, liking, and
tagging. Furthermore, Facebook allows user to interact with entities
(groups, pages and favourites) via membership and liking. The app also
keep tracks of the information of user's group membership, page likes
and favourites.  The App does not track deletions of these items,
interactions (e.g., unlike) and memberships for performance reasons
and we found very few deletions during an initial testing stage.  We
summarize relevant basic statistics of the data in
Table~\ref{tab:interactions}-\ref{tab:likeinfo} below.
Table~\ref{tab:interactions} summarizes the number of records for each
item (row) and interaction (column)
combination. Table~\ref{tab:demographics} shows some demographics from
user profiles. Table \ref{tab:interests} shows the group membership,
page like and favourites counts for app users and all users.

Our app recommends links to users every day, where the users may give
their feedback on the links indicating whether they liked it or
disliked it. Users are recommended links that are collected either
from friends or non-friends. Throughout this paper, app user rated
link data is used to evaluate the recommendation algorithms.
Table \ref{tab:likeinfo} shows the statistics of the app user rating
for friend and non-friend recommendations.
      							
\begin{table}
\centering
\begin{tabular}{|>{\small}l|>{\small}r|>{\small}r|}
\hline
& \textbf{App Users} & \textbf{Ego network} \\
& & \textbf{of App Users} \\
\hline
Users & 119 & 38,378 \\
\hline
Male & 85 & 20,840 \\
\hline
Female & 33 & 17,032 \\
\hline
\end{tabular}
\caption{App user demographics}
\label{tab:demographics}
\end{table}

\begin{table}
\centering
\begin{tabular}{|>{\small}l|>{\small}r|>{\small}r|>{\small}r|}
\hline
\textbf{App Users} & \textbf{Tags} & \textbf{Comments} & \textbf{Likes} \\
\hline
\textbf{Post} & 7,711 & 22,388 & 15,999 \\
\hline
\textbf{Link}  & --- & 7,483 & 6,566 \\
\hline
\textbf{Photo} & 28,341 & 10,976 & 8,612 \\
\hline
\textbf{Video} & 2,525 & 1,970 & 843 \\
\hline
\hline
\textbf{Ego network} & \textbf{Tags} & \textbf{Comments} & \textbf{Likes} \\
\textbf{of App Users}  & & & \\
\hline
\textbf{Post} & 1,215,382 & 3,122,019 & 1,887,497 \\
\hline
\textbf{Link} & --- & 891,986 & 995,214 \\
\hline
\textbf{Photo} & 9,620,708 & 3,431,321 & 2,469,859 \\
\hline
\textbf{Video} & 904,604 & 486,677 & 332,619 \\
\hline
\end{tabular}
\caption{Statistics on user {\em Interactions}, grouped by item {\em Modality} and {\em Action type}.}
\label{tab:interactions}
\end{table}


\begin{table}[t!]
\centering
\begin{tabular}{|>{\small}l|>{\small}r|>{\small}r|}
\hline
& \textbf{App Users} & \textbf{Ego Network} \\
& & \textbf{of App Users} \\
\hline
Groups & 3,469 & 373,608 \\
\hline
Page Likes & 10,771 & 825,452 \\
\hline
Favourites & 4,284 & 892,820\\
\hline
\end{tabular}
\caption{Statistics on user {\em Actions}, counted for {\em groups, pages} and {\em favorites} over the App users and their ego network.}
\label{tab:interests}
\end{table}

\begin{table}[t!]
\centering
\begin{tabular}{|>{\small}l|>{\small}r|>{\small}r|}\hline
&\textbf{Friend}  & \textbf{Non-Friend} \\
&\textbf{recommendation}  & \textbf{recommendation} \\
\hline
Like& 1392 & 1127 \\
\hline
Dislike& 895 & 2111\\
\hline
\end{tabular}
\caption{Dataset breakdown of prediction target $like(u,i)$, by the source of the link (friend/non-friend) and value (true/false).}
\label{tab:likeinfo}
\end{table}


\eat{ %% suvash's original table
\begin{table}
\centering
\begin{tabular}{|>{\small}l|>{\small}r|>{\small}r|>{\small}r|>{\small}r|}
\hline
\textbf{App Users} & \textbf{Posts} & \textbf{Tags} & \textbf{Comments} & \textbf{Likes} \\
\hline
\textbf{Wall} & 36,359 & 7,711 & 22,388 & 15,999 \\
\hline
\textbf{Link} & 5,304 & --- & 7,483 & 6,566 \\
\hline
\textbf{Photo} & 4,933 & 28,341 & 10,976 & 8,612 \\
\hline
\textbf{Video} & 245 & 2,525 & 1,970 & 843 \\
\hline
\hline
\textbf{App Users} & \textbf{Posts} & \textbf{Tags} & \textbf{Comments} & \textbf{Likes} \\
\textbf{and Friends} & & & & \\
\hline
\textbf{Wall} & 4,301,306 & 1,215,382 & 3,122,019 & 1,887,497 \\
\hline
\textbf{Link} & 678,612 & --- & 891,986 & 995,214 \\
\hline
\textbf{Photo} & 1,268,816 & 9,620,708 & 3,431,321 & 2,469,859 \\
\hline
\textbf{Video} & 59,244 & 904,604 & 486,677 & 332,619 \\
\hline
\end{tabular}
\caption{The number of interactions Tables. Rows are type of Facebook item and columns are type of Facebook interaction.}
\label{tab:interactions}
\end{table}
}
