%!TEX root = document.tex

% SPS: Don't really need a conclusions section... the intro section
% already did this.
\eat{
We proposed Social Affinity Filtering (SAF) as a novel method for social
recommendation that analyses a user's fine-grained interactions and
activities to learn which subset of social groups have the highest affinity with
a user's preferences.  We evaluated SAF on a dataset collected from a
Facebook App, showing that SAF yields 6\% absolute improvement in
accuracy over state-of-the-art (social) recommendation engines just
using knowledge of users' page likes.  This is an important result
given that SAF is built on standard supervised linear classification
techniques (which support strong training guarantees and fast learning
algorithms) and would be fairly robust in an online setting 
in contrast to some more complex matrix factorization optimisation approaches
often proposed for (social) collaborative filtering.

In addition to the strong social recommendation accuracy improvements
offered by SAF, we quantified the relative importance of interaction
and activity groups for recommendation and we analysed what properties
made some social affinity groups more informative than others.  Among
many insights, our results show that video interactions are more
predictive than other modalities and outgoing interactions are more
predictive than incoming ones, but both can be superceded by the
number of preferences expressed by a group.  Furthermore, for
activities, we showed that smaller social groups are more predictive
than larger ones and long-tailed categories with many specialised
choices (dynamically increasing over time) tend to contain some of
the most predictive affinity groups.  Knowing what subset of features
are informative allows one to design an effective recommendation tool
that requires minimal permissions from a user --- a key property for
general user uptake.
}

For a summary of the key contributions and insights from this work, we
refer the reader back to the points outlined in
Sec~\ref{sec:introduction}.  Future directions of research along the
lines of SAF can examine the nature of social groups via additional
metrics --- e.g., the social network within members of the group, or
activity level of the group.  Other work might explore non-binary
representations of ASAFs and ISAFs that enhance the predictive power
of SAF, or extensions of (social) collaborative filtering approaches
like nearest neighbor or matrix factorization to better exploit
fine-grained social interaction and activity features as in SAF.

%% Latent projections of affinity groups!  (for social regularization)
%% Exploit quantity in features and interactions/nonlinearity
%% Cold-start

%% Explore beyond one link out... two links, three links... how?
%% Try latent factorization approach?
%% Taxonomic features?
