%!TEX root = document.tex

We built a Facebook App to collect explicit user like and dislike 
preferences back for links posted on Facebook (e.g., Youtube video,
news or blog item, etc.) as well as detailed user interaction and
activity history available through the Facebook Graph API.
%The purpose of the App was two-fold: (1) to build the SAG information
%as defined in Sec~\ref{ssec:sag}, and (2) to collect a user $u$'s
%preferences on recommended links $i$ solicited by the App on a daily
%basis to build the $\likes(u,i)$ relation required in
%Sec~\ref{ssec:SAfeature}.
The data collection was performed with full permission from the user
and in accordance with an approved Ethics Protocol 
\#2011/142 from the Australian National University.

Our App requested to collect information on profiles (including
activity memberships) and timelines (interactions) for the App
users \emph{and} their friends as required by Sec~\ref{ssec:sag} and
Sec~\ref{ssec:SAfeature}.  With such expressive permissions, many
potential users were hesitant to install the App --- after an
intensive one month user drive at our University, we were able to
attract 119 App users allowing us to collect activity and interaction
data for a combined 37,872 users.\footnote{The issue of low App user
uptake with such expressive App permissions underscores the importance
of identifying the \emph{minimal} set of permissions to obtain good
recommendation performance --- a question we address in our subsequent
analysis.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{table}[t!]
\centering
\begin{tabular}{|>{\small}l|>{\small}r|>{\small}r|}
\hline
& \textbf{App Users} & \textbf{Ego network} \\
& & \textbf{of App Users} \\
\hline
Total & 119 & 37,872 \\
\hline \hline
Male & 85 & 20,840 \\
\hline
Female & 34 & 17,032 \\
\hline
\end{tabular}
\caption{App user demographics.  The \emph{ego network} is the friend
network of the App users.}
\label{tab:demographics}
\end{table}

\begin{table}[t!]
\centering
\begin{tabular}{|>{\small}l|>{\small}r|>{\small}r|>{\small}r|}
\hline
\textbf{App Users} & \textbf{Tags} & \textbf{Comments} & \textbf{Likes} \\
\hline
\textbf{Post} & 7,711 & 22,388 & 15,999 \\
\hline
\textbf{Link}  & --- & 7,483 & 6,566 \\
\hline
\textbf{Photo} & 28,341 & 10,976 & 8,612 \\
\hline
\textbf{Video} & 2,525 & 1,970 & 843 \\
\hline
\hline
\textbf{Ego network} & \textbf{Tags} & \textbf{Comments} & \textbf{Likes} \\
\textbf{of App Users}  & & & \\
\hline
\textbf{Post} & 1,215,382 & 3,122,019 & 1,887,497 \\
\hline
\textbf{Link} & --- & 891,986 & 995,214 \\
\hline
\textbf{Photo} & 9,620,708 & 3,431,321 & 2,469,859 \\
\hline
\textbf{Video} & 904,604 & 486,677 & 332,619 \\
\hline
\end{tabular}
\caption{Statistics on user {\em interactions}.
%grouped by item {\em modality} and {\em action type}.
}
\label{tab:interactions}
\end{table}

\begin{table}[t!]
\centering
\begin{tabular}{|>{\small}l|>{\small}r|>{\small}r|}
\hline
& \textbf{App Users} & \textbf{Ego Network} \\
& & \textbf{of App Users} \\
\hline
Groups & 3,469 & 373,608 \\
\hline
Page Likes & 10,771 & 825,452 \\
\hline
Favourites & 4,284 & 892,820\\
\hline
\end{tabular}
\caption{Statistics on user {\em actions}, counted for {\em Groups, Pages} and {\em Favourites} over the App users and their ego network.}
\label{tab:interests}
\end{table}

\begin{table}[t!]
\centering
\begin{tabular}{|>{\small}l|>{\small}r|>{\small}r|}\hline
&\textbf{Friend}  & \textbf{Non-Friend} \\
&\textbf{recommendation}  & \textbf{recommendation} \\
\hline
Like& 1392 & 1127 \\
\hline
Dislike& 895 & 2111\\
\hline
\end{tabular}
\caption{Dataset breakdown of prediction target $like(u,i)$ by the source of the link (Friend/Non-friend) and rating (Like=$\true$, Dislike=$\false$).}
\label{tab:likeinfo}
\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

We summarise basic statistics of the data in
Tables~\ref{tab:demographics}--\ref{tab:likeinfo}.
Table~\ref{tab:demographics} presents user and friend demographics.
Table~\ref{tab:interactions} summarises the number of
records for each item modality (row) and action (column)
combination. Table~\ref{tab:interests} shows the group membership,
page like and favourite counts for users.

Our App recommends three links to App users each day,
which the users may optionally like or dislike.  Recommended links are
harvested from
\emph{both} friends' and non-friends' timelines.
We display only three links per day in order to avoid rank-bias with
preferences; each link could be independently rated.
Table~\ref{tab:likeinfo} shows App user
link preference statistics.

All subsequent experimentats use offline
batch data stored and analysed \emph{after} a four
month data collection period.
      							
