\chapter*{Abstract}
\addcontentsline{toc}{chapter}{Abstract}

Automatic recommendations helps users see what items may be important them, and saves them the time and effort of looking for that item. The most basic recommender systems make recommendation using information about the users themselves as well as the items that are being recommended. Collaborative recommendation extends this by basing recommendations on correlations between other users with the same preferences. This paper deals with another extension to collaborative recommendation systems, that of social recommendation. Social recommendation makes use of a user's social information to improve the recommendations made to those users. Social information can come from the data in social networks like Facebook like the user's friends on the network or a $\log$ of the normalized sum of interactions between users. This paper looks at different methods of incorporating social information into recommender systems, in particular link recommendation on Facebook. Two sets of live user trial was done, using the LinkR Facebook application that was developed as part of this project  Offline experiments were also performed on Facebook data that gathered by LinkR. Lastly, a user survey was conducted at the end of the first  user trial and those results are also reported here. The results of the the first user trial, offline experiments, and the user survey were used to design new methods for social algorithms and a second live user trial was done with the new algorithms.