\chapter*{Abstract}
\addcontentsline{toc}{chapter}{Abstract}

\begin{comment}
Automatic recommendations helps users see what items may be important
them, and saves them the time and effort of looking for that item. The
most basic recommender systems make recommendation using information
about the users themselves as well as the items that are being
recommended. Collaborative recommendation extends this by basing
recommendations on correlations between other users with the same
preferences. This thesis deals with another extension to collaborative
recommendation systems, that of social recommendation. Social
recommendation makes use of a user's social information to improve the
recommendations made to those users. Social information can come from
the data in social networks like Facebook like the user's friends on
the network or a $\log$ of the normalized sum of interactions between
users. This thesis looks at different methods of incorporating social
information into recommender systems, in particular link
recommendation on Facebook. Two sets of live user trial was done,
using the LinkR Facebook application that was developed as part of
this project Offline experiments were also performed on Facebook data
that gathered by LinkR. Lastly, a user survey was conducted at the end
of the first user trial and those results are also reported here. The
results of the the first user trial, offline experiments, and the user
survey were used to design new methods for social algorithms and a
second live user trial was done with the new algorithms.
\end{comment}

This thesis examines the problem of designing efficient, scalable, and
accurate social \emph{collaborative filtering} (CF) algorithms for
personalized link recommendation on Facebook.  Unlike standard CF
algorithms using relatively simple user and item features (possibly
just the user ID and link ID), link recommendation on social networks
like Facebook poses the more complex problem of learning user
preferences from a rich and complex set of user profile and
interaction information.  Most existing \emph{social CF} (SCF) methods
have extended traditional CF \emph{matrix factorization} (MF)
approaches, but have overlooked important aspects specific to the
social setting; specifically, existing SCF MF methods (a) do not
permit the use of item or link features in learning user similarity
based on observed interactions, (b) do not permit directly modeling
user-user information diffusion according to the social graph
structure, and (c) cannot learn that that two users may only have
overlapping interests in specific areas.  
This thesis proposes a unified SCF optimization framework that
addresses (a)--(c) and compares these novel algorithms with a variety
of existing baslines.  Evaluation is carried out via live user trials
in a custom-developed Facebook App involving data collected over three
months from over 100 App users and their nearly 30,000 friends.  Not
only do we show that our novel proposals to address (a)--(c)
outperform existing approaches, but we also identify which offline
ranking and classigication evaluation metrics correlate most with
human judgment of algorithm performance.  Overall, this thesis
represents a critical step forward in extending SCF recommendation
algorithms to fully exploit the rich content and structure of social
networks like Facebook.